Extracting Information from AGN Variability

### 2016-05-20 Manuscript.tex
Text done...just need to replot

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[V.~P. Kasliwal et al.]{
\parbox{\textwidth}{
Vishal~P. Kasliwal,$^{1,2}$\thanks{E-mail: \texttt{vish@sas.upenn.edu}}
Michael~S. Vogeley,$^{3}$
Gordon~T. Richards$^{3}$
\& Nate~B. Lust$^{2}$}\vspace{0.4cm}\
\\
% List of institutions
\parbox{\textwidth}{
$^{1}$Department of Physics \& Astronomy, University of Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104-6396, USA\\
$^{2}$Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ, 08544, USA\\
$^{3}$Department of Physics, Drexel University, Philadelphia, PA 19104-2875, USA\\
%$^{2}$Department of Physics and Astronomy, Western Kentucky University, Bowling Green, KY 42101-1077, USA
}
}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2015}

@@

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variability across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood because of the complexity of the phenomenon being observed. While AGN variability is usually viewed as useful for AGN detection, there is a lack of sophisticated, non-power spectral density (PSD) based techniques to use variability as a probe of accretion physics. We present a new method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the PSD of the driving impulses using the C-ARMA model of \citet{Kelly14}.

We decompose the observed fluctuations in light curves into stochastic impulses that drive an ordinary linear differential equation. The Green's function of the differential equation quantifies the time-domain evolution of perturbations to the emitted flux and hence probes dissipative physical processes in accretion flow. The PSD of the stochastic driving impulses characterizes the physical processes responsible for the flux perturbations. We present a new state-space representation of the C-ARMA process of \citet{Kelly14} that is better suited for analyzing regularly sampled light curves with missing observations. Extension of the C-ARMA process to more sophisticated models is also easier in this representation.

Applying our technique to the light curve of the \Kepler AGN Zw 229-15, we find that the observed variability behavior is consistent with that of a damped harmonic oscillator driven by a colored noise process. The model PSD turns over on time-scale $385$~d to a flat 'white noise' spectrum. On shorter time-scales, the model PSD log slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify the $5.6$~d time-scale reported by \citet{Edelson14} as the time-scale on which flux perturbations grow. We also recover and identify the $67$~d time-scale reported by \citet{CariniWilliamsAAS} with the time-scale on which flux perturbations decay back to the steady-state flux level. The PSD of the driving impulses has a flat 'white noise' power spectrum on long time-scales. On time-scales shorter than $1$~d, the power spectrum begins to rise steeply, characteristic of 'violet noise' spectra. The driving impulse PSD may be caused by turbulent eddies in the accretion disk, or by the thermal motion of the accretion disk plasma. We conclude that valuable physical insights into accretion may be obtained using our method. 
\end{abstract}

@@

Active Galactic Nuclei (AGN) exhibit variability on time-scales ranging from minutes and hours to years over the full electromagnetic spectrum. The underlying cause of the variability is not clear. Proposed models range from local viscosity fluctuations to oscillatory modes in the accretion disk \citep{UrryARAA}. AGN are powered by the accretion of matter onto a central supermassive black-hole \citep{ReesARAA}. Matter infalling onto the supermassive black hole must lose angular momentum during inspiral. The loss of angular momentum results in the conversion of gravitational binding energy to the kinetic energy of the flow. A portion of this kinetic energy is radiated away resulting in the characteristic appearance of the AGN--high luminosity ($\sim 10^{47}$~erg~s$^{-1}$ for the brightest quasars) emanating from a small volume ($\sim 10^{14}$~cm) of space \citep{Edelson96}. Open questions include: (1) What is the angular momentum-loss mechanism responsible for creating an inward flow? and (2) What is the process responsible for the highly efficient conversion of the kinetic energy of the accretion flow into radiation?

It is generally accepted that the accretion inflow results in the formation of an accretion disk \citep*{KoratkarBlaes99,PringleARAA}. At least three mechanisms may be responsible for the extraction of angular momentum from the accretion flow in a standard disk: (1) Magneto-rotational Instability (MRI) generated turbulence \citep*{BHI,BHII,BHIII,BHIV,BHG95,BHRev}; (2) Large scale magneto-hydrodynamic (MHD) outflow caused external stresses \citep*{BlandfordPayne82}; and (3) Shocks produced by non-axis-symmetric waves \citep{FragileBlaes08}. MRI prescriptions for angular momentum transport are succinctly quantified by the alpha-prescription of \citet{ShakuraSunyaev73}--see \citet{BalbusPapaloizou99}. Under the alpha prescription, accretion flows are modelled as solutions of the vertically integrated hydrodynamic equations for the conservation of mass, radial and angular momentum, and energy. Stability criteria then dictate that solutions be of one of three types as characterized by location in surface density-mass accretion rate phase space \citep{BlaesAccretion}. Low accretion rate coupled with low surface density results in the formation of advection-dominated accretion flows \citep{NarayanYi94,NarayanYi95a,NarayanYi95b,AbramowiczChen95,ChenAbramowicz95} where the cooling time of the accretion flow is much greater than the in-fall time. Such disks are optically thin and may exist in low-luminosity AGN. Some quasars, a few Seyfert 1s and the intermediate states of black-hole X-ray binaries (BHXRB) are thought to possess radiation pressure dominated 'slim' disks ($H/r < 1$ where $H$ is the local disk height at radius $r$) with medium surface mass densities and very high accretion rates \citep{Abramowicz88}. Such disks also have shorter inflow time as compared to the cooling time-scale making the flow advective. The third stable solution occurs at high surface mass density but low accretion rate and results in the formation of classic 'thin' accretion disks that are optically thick and geometrically thin ($H/r << 1$) \citep*{ShakuraSunyaev73,FKR}. Most AGN and BHXRBs in the high and soft states are thought to possess such disks \citep{BlaesAccretion}.

@@

AGN optical variability may arise due to fluctuations in the \citet{ShakuraSunyaev73} viscosity parameter $\alpha$. \citet{Lyubarskii97} showed that local stochastic fluctuations in the accretion disk are capable of producing the power-law power spectral densities (PSD) observed in AGN and BHXRBs. Such fluctuations are thought to propagate inwards over time suggesting that a powerful observational test is the observation of blue-lags \citep{UttleyAccretion} i.e. light curve variations at shorter wavelengths lagging those at longer wavelengths. Evidence for the propagating fluctuation model was provided by \citet{Miyamoto88} who observed that hard X-ray variations in Cyg X-1 lagged behind soft X-ray variations with an inverse scaling between the time lag and the variability time-scale. \citet{Starling04} used the viscosity fluctuation model to impose lower limit of $\alpha \sim 0.01$ on the disk viscosity. A later study by \citet{King07} using AGN variability triggered by viscosity fluctuations suggest that $\alpha \sim 0.1$-$0.4$. \citet{Wood01} examined a variant of the \citet{Lyubarskii97} propagating fluctuations model in which episodic mass deposition at the outer edge of the accretion disk triggers viscosity fluctuations in the accretion disk that evolve via the standard non-linear diffusion equation for the surface mass density
\begin{equation}\label{eq:SurfaceMassDensity}
\frac{\partial \Sigma}{\partial t} = \frac{3}{r} \frac{\partial}{\partial r} \left[ \sqrt{r} \frac{\partial \left( \nu \Sigma \sqrt{r}  \right)}{\partial r} \right],
\end{equation}
where $\nu = \alpha c_{S} H$ is the kinematic viscosity and $\Sigma$ is the surface mass density \citep{PringleARAA,FKR}. This model was revisited by \citet{Titarchuk07} who generalized the treatment of \citet{Wood01} to include time-dependent fluctuation sources located randomly on the disk face. By applying the surface mass density equation~\eqref{eq:SurfaceMassDensity}, \citet{Titarchuk07} follow the evolution of perturbations to the mass surface density and luminosity arising from viscosity that is a power-law function of disk radius. Subsequently, \citet{Kelly11} studied solutions to the surface mass density equation perturbed by stochastic fluctuations and showed that the resulting light curve can be expressed as a linear combination of multiple Ornstein-Uhlenbeck processes \citep{Kelly09,Gillespie96}.

Accretion disk variability may also be caused by the existence of `hot-spots' in the accretion disk \citep{MaccaroneAccretion}. Hot spots are thought to arise when accretion streams impact the disk resulting in shock heating. Such hot spots have already been detected during periods of quiescence in BHXRBs \citet{Froning11,McClintock95}. Alternatively, thermal viscous instabilities or a variable mass transfer rate may be responsible for accretion disk variability \citep{Lasota01,Coriat12}, though the amplitude of the variability should be much smaller and the time-scale much longer for AGN than for BHXRBs \citep{Hameury09}.

Other sources of variability include magnetic flares in the corona of the accretion disk \citep{PoutanenFabian99} or from the hot accretion flow itself \citep{Veledina13}. Alternatively, dynamo processes due to the magnetic fields threading the disk may be responsible for producing variability \citep{LivioPringleKing03,King04,MayerPringle06}. \citet{JaniukCzerny07} have suggested that the variability may be produced in the hot X-ray corona surrounding the accretion disk from the evaporation of mass perturbations in the disk. \citet{Misra08} suggested that the X-ray variability of Cyg X-1 can be modelled as a damped harmonic oscillator driven by noise. Quasi-periodic signals in the light curve may be produced by weak shocks in the innermost region of the accretion caused by the presence of a tilt in the accretion disk of the black hole \citep{FragileBlaes07,FragileBlaes08}. Alternatively, quasi-periodic signals may be caused by global oscillation modes or parametric resonance instability \citep{ReynoldsMiller09a,ReynoldsMiller09b,Oneill11}. Variability may also be produced by standing shocks interacting with fluctuations in tilted accretion disks \citep{Henisey12}.

@@

Several mechanisms may be responsible for the variability observed in AGN light curves. Light curves in the optical are driven by X-ray variability on shorter time-scales but exhibit a local source of variability on longer time-scales \citep{UttleyAccretion}. This implies that it is unlikely that a single mechanism drives variability across the entire electromagnetic spectrum. Rather, the nature of the underlying mechanism responsible for variability may be different at different wavelengths.

@@

We propose a Green's function-based method for analysing observed light curves, particularly in the UV through IR where the fluctuations are smaller than they tend to be in the X-ray, that facilitates comparison with analytic and simulation work on variability. Our method builds on the Continuous-time AutoRegressive Moving Average (C-ARMA) modelling technique of \citet{Kelly14} by proposing that the C-ARMA process be interpreted literally as quantifying the physics governing variability i.e. the C-ARMA process is a direct result of the physics driving the variability.

As discussed in the introduction, it is likely that variability originates from multiple mechanisms simultaneously. These mechanisms are almost certainly non-linear \citep{UMV05}. We begin by linearising a general non-linear differential equation such as the surface mass density equation of~\eqref{eq:SurfaceMassDensity} that governs how perturbations evolve. We shall find that the linearisation of the non-linear equation results in a linear differential equation. This linearised equation has the form of the left hand side (LHS) of the C-ARMA process of equation~\eqref{eq:CARMAIntro}. Since the C-ARMA process is linear, we suggest that the Green's function of the LHS of equation~\eqref{eq:CARMAIntro} be used to probe dissipation physics in the accretion disk.

The right hand side (RHS) of the C-ARMA process of equation~\eqref{eq:CARMAIntro} can then be viewed as driving the LHS, i.e. a C-ARMA process is a linear differential equation driven by a random process. This driving process may be viewed as disturbances generated by MRI instabilities in the accretion disk. Section~\ref{sec:RHS} proposes a form for the disturbances driving the flux perturbations: coloured noise stochastic processes. A simple example of a C-ARMA process--the C-ARMA($2$,$1$) process--is presented in section~\ref{sec:CARMA21}.

@@

We assume that the variations in AGN luminosity may be ascribed to flux perturbations intrinsic to the accretion disk, corona and winds. Rather than assuming an underlying mechanism for the flux perturbations, we shall only assume that the perturbations are generated by a complex physical process. We assume that the total flux emitted by the AGN evolves via a (probably non-linear) differential equation. We shall examine how linear perturbations of the total flux evolve. Let the total flux emitted by an AGN be $F(t)$. Suppose $F$ obeys the $p^{\mathrm{th}}$-order non-linear differential equation 

@@

where $\mathrm{d}u$ is an increment of $u$. We see that the LHS of this equation is identical to the LHS of the C-ARMA process equation~\eqref{eq:CARMAIntro} suggesting that a C-ARMA process may be a good model for small flux variations.

@@

Thus, we find that if the total flux emitted by the accretion disk obeys a non-linear differential equation, we can linearize that equation to determine the behaviour of small perturbations of the flux from the steady-state solution. The linear differential equation that governs how a small flux perturbation evolves is identical to the C-ARMA process equation suggesting that such processes may be good models for the stochastic variability seen in AGN accretion disks. In the next section, we present a convenient physical interpretation of equation~\eqref{eq:FinalSystemLHS} using the method of the Green's function.

@@

We compute the Green's function of equation~\eqref{eq:FinalSystemLHS} by finding a solution of the homogenized version of the equation. Consider the homogeneous differential equation corresponding to equation~\eqref{eq:FinalSystemLHS}
\begin{equation}\label{eq:HomoCARMA}
\frac{\mathrm{d}^{p}f}{\mathrm{d}t^{p}} + \alpha_{1} \frac{\mathrm{d}^{p-1}f}{\mathrm{d}t^{p-1}} + \ldots + \alpha_{p-1} \frac{\mathrm{d}f}{\mathrm{d}t} + \alpha_{p} f = 0.
\end{equation}
To obtain the Green's function solution of this equation, we shall drive this equation with a unit impulse, i.e. a Dirac delta function, located at $t = 0$
\begin{equation}\label{eq:GreenCARMA}
\frac{\mathrm{d}^{p}G}{\mathrm{d}t^{p}} + \alpha_{1} \frac{\mathrm{d}^{p-1}G}{\mathrm{d}t^{p-1}} + \ldots + \alpha_{p-1} \frac{\mathrm{d}G}{\mathrm{d}t} + \alpha_{p} G = \delta(t),
\end{equation}
where $G(t)$ is the desired Green's function solution. To find this solution, we shall first solve equation~\eqref{eq:HomoCARMA} by finding the roots of the characteristic polynomial of equation~\eqref{eq:ARCharPoly} and then apply boundary conditions based on the properties of $\delta(x)$. From the definition of the characteristic polynomial in equation~\eqref{eq:ARCharPoly}, the (distinct) roots are $\rho_{k}$ with $\operatorname{Re}(\rho_{k}) < 0$ for stability. Then the solution of the homogeneous equation~\eqref{eq:HomoCARMA} is 
\begin{equation}\label{eq:HomoSol}
f(t) = \sum_{k = 1}^{n} c_{k} \mathrm{e}^{\rho_{k}t},
\end{equation}
where the $c_{k}$ are constants. Hence the solution of the inhomogeneous equation~\eqref{eq:GreenCARMA} may be found by using the continuity properties of $G(t)$ and its derivatives. For $t < 0$, $\mathrm{d}^{p-k}G/\mathrm{d}t^{p-k} = 0$ for $1 \leq k < p$ while at $t = 0$, $\mathrm{d}^{p}G/\mathrm{d}t^{p}$ has the same type of discontinuity as $\delta(t)$. Therefore $\mathrm{d}^{p-1}G/\mathrm{d}t^{p-1} = 1$ when $t = 0$, i.e. it behaves like the step function. Next, we have that $\mathrm{d}^{p-k}G/\mathrm{d}t^{p-k} = 0$ at $t = 0$ for all $1 < k \leq p$ since each successive lower-order derivative is a $(k-1)^{\mathrm{th}}$-order polynomial located at the origin. Following this line of reasoning, we arrive at the following boundary conditions
\begin{equation}\label{eq:BC1a}
 \left.\frac{\mathrm{d}^{p-1}G}{\mathrm{d}t^{p-1}}\right|_{t = 0} = 1,\\
\end{equation}
and
\begin{equation}\label{eq:BC2a}
 \left.\frac{\mathrm{d}^{p-k}G}{\mathrm{d}t^{p-k}}\right|_{t = 0} = 0,
\end{equation}
for all $2 \leq k \leq p$. Computation of the derivatives in equations~\eqref{eq:BC1a} and \eqref{eq:BC2a} results in
\begin{equation}\label{eq:BC1b}
 \sum_{k = 1}^{p} c_{k} \rho_{k}^{p-1} = 1,
 \end{equation}
 and
\begin{equation}\label{eq:BC2b}
 \sum_{k = 1}^{p} c_{k} \rho_{k}^{k-1} = 0,
\end{equation}
for all $2 \leq k \leq p$. The $c_{k}$ may then be found by solving the matrix equation
\begin{equation}\label{eq:BCs}
\left( \begin{array}{ccccc} 
\rho_{1}^{p-1} & \rho_{2}^{p-1} & \hdots & \rho_{p-1}^{p-1} & \rho_{p}^{p-1} \\
\rho_{1}^{p-2} & \rho_{2}^{p-2} & \hdots & \rho_{p-1}^{p-2} & \rho_{p}^{p-2} \\
\vdots & \vdots & \ddots & \vdots & \vdots \\
\rho_{1} & \rho_{2} & \hdots & \rho_{p-1} & \rho_{p} \\
1 & 1 & \hdots & 1 & 1 \\
\end{array}\right) \left( \begin{array}{c} c_{1} \\ c_{2} \\ \vdots \\ c_{p-1} \\ c_{p} \end{array} \right) = \left( \begin{array}{c} 1 \\ 0 \\ \vdots \\ 0 \\ 0 \end{array} \right).
\end{equation}
To summarize, the Green's function for equation~\eqref{eq:FinalSystemLHS} is given by 
\begin{equation}\label{eq:GreenSol}
G(t) = \sum_{k = 1}^{p} c_{k} \mathrm{e}^{\rho_{k}t},
\end{equation}
where the $c_{k}$ satisfy equation~\eqref{eq:BCs}.

\section[Driving Perturbations]{What Drives the Flux Perturbations?}\label{sec:RHS}

@@

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[V.~P. Kasliwal et al.]{
\parbox{\textwidth}{
Vishal~P. Kasliwal,$^{1,2}$\thanks{E-mail: \texttt{vish@sas.upenn.edu}}
Michael~S. Vogeley,$^{3}$
\&Gordon~T. Richards$^{3}$}\vspace{0.4cm}\
\\
% List of institutions
\parbox{\textwidth}{
$^{1}$Department of Physics \& Astronomy, University of Pennsylvania, 209 South 33rd Street, Philadelphia, PA 19104-6396, USA\\
$^{2}$Department of Astrophysical Sciences, 4 Ivy Lane, Princeton University, Princeton, NJ, 08544, USA\\
$^{3}$Department of Physics, Drexel University, Philadelphia, PA 19104-2875, USA\\
%$^{2}$Department of Physics and Astronomy, Western Kentucky University, Bowling Green, KY 42101-1077, USA
}
}

@@

% Enter the current year, for the copyright statements etc.
\pubyear{2016}

@@

% Abstract of the paper
\begin{abstract}
AGN exhibit rapid, high amplitude stochastic flux variability across the entire electromagnetic spectrum on timescales ranging from hours to years. The cause of this variability is poorly understood because of the complexity of the phenomenon being observed. We present a new method for using variability to (1) measure the time-scales on which flux perturbations evolve and (2) characterize the driving flux perturbations. We decompose the observed light curve fluctuations into stochastic impulses that drive an ordinary linear differential equation. The Green's function of the differential equation quantifies the time-domain evolution of perturbations to the emitted flux and hence probes dissipative physical processes in accretion flow. The PSD of the stochastic driving impulses characterizes the physical processes responsible for the flux perturbations. Applying our technique to the light curve of the \Kepler AGN Zw 229-15, we find that the observed variability behavior can be modeled as a damped harmonic oscillator perturbed by a colored noise process. The model PSD turns over on time-scale $385$~d. On shorter time-scales, the model PSD log slope varies between $2$ and $4$, explaining the behavior noted by previous studies. We recover and identify both the $5.6$~d and $67$~d timescales reported by previous work. In our model, these timescales represent the time-scale on which flux perturbations grow, and the time-scale on which flux perturbations decay back to the steady-state flux level respectively. The PSD of the driving impulses has a flat power spectrum on long time-scales. On time-scales shorter than $1$~d, the power spectrum begins to rise steeply possibly due to turbulent eddies in the accretion disk, or by the thermal motion of the accretion disk plasma. We conclude that valuable physical insights into accretion may be obtained using our method.
\end{abstract}

@@

Active Galactic Nuclei (AGN) exhibit variability on time-scales ranging from minutes and hours to years over the full electromagnetic spectrum. The underlying cause of the variability is not clear. Proposed models range from local viscosity fluctuations to oscillatory modes in the accretion disk \citep{UrryARAA}. AGN are powered by the accretion of matter onto a central supermassive black-hole \citep{ReesARAA}. Matter infalling onto the supermassive black hole must lose angular momentum during inspiral. The loss of angular momentum results in the conversion of gravitational binding energy to the kinetic energy of the flow. A portion of this kinetic energy is radiated away resulting in the characteristic appearance of the AGN--high luminosity ($\sim 10^{47}$~erg~s$^{-1}$ for the brightest quasars) emanating from a small volume ($\sim 10^{14}$~cm) of space \citep{Edelson96}.

It is generally accepted that the accretion inflow results in the formation of an accretion disk \citep*{KoratkarBlaes99,PringleARAA}. At least three mechanisms may be responsible for the extraction of angular momentum from the accretion flow in a standard disk: (1) Magneto-rotational Instability (MRI) generated turbulence \citep*{BHI,BHRev}; (2) Large scale magneto-hydrodynamic (MHD) outflow caused external stresses \citep*{BlandfordPayne82}; and (3) Shocks produced by non-axis-symmetric waves \citep{FragileBlaes08}. MRI prescriptions for angular momentum transport are succinctly quantified by the alpha-prescription of \citet{ShakuraSunyaev73}--see \citet{BalbusPapaloizou99}. Under the alpha prescription, accretion flows are modelled as solutions of the vertically integrated hydrodynamic equations for the conservation of mass, radial and angular momentum, and energy. Stability criteria then dictate that solutions be of one of three types as characterized by location in surface density-mass accretion rate phase space \citep{BlaesAccretion}. Low accretion rate coupled with low surface density results in the formation of advection-dominated accretion flows \citep{NarayanYi94,ChenAbramowicz95} where the cooling time of the accretion flow is much greater than the in-fall time. Such disks are optically thin and may exist in low-luminosity AGN. Some quasars, a few Seyfert 1s and the intermediate states of black-hole X-ray binaries (BHXRB) are thought to possess radiation pressure dominated 'slim' disks ($H/r < 1$ where $H$ is the local disk height at radius $r$) with medium surface mass densities and very high accretion rates \citep{Abramowicz88}. Such disks also have shorter inflow time as compared to the cooling time-scale making the flow advective. The third stable solution occurs at high surface mass density but low accretion rate and results in the formation of classic 'thin' accretion disks that are optically thick and geometrically thin ($H/r << 1$) \citep*{ShakuraSunyaev73,FKR}. Most AGN and BHXRBs in the high and soft states are thought to possess such disks \citep{BlaesAccretion}.

@@

AGN optical variability may arise due to fluctuations in the \citet{ShakuraSunyaev73} viscosity parameter $\alpha$. \citet{Lyubarskii97} showed that local stochastic fluctuations in the accretion disk are capable of producing the power-law power spectral densities (PSD) observed in AGN and BHXRBs. Such fluctuations are thought to propagate inwards over time suggesting that a powerful observational test is the observation of blue-lags \citep{UttleyAccretion} i.e. light curve variations at shorter wavelengths lagging those at longer wavelengths. Evidence for the propagating fluctuation model was provided by \citet{Miyamoto88} who observed that hard X-ray variations in Cyg X-1 lagged behind soft X-ray variations with an inverse scaling between the time lag and the variability time-scale. \citet{Starling04} used the viscosity fluctuation model to impose lower limit of $\alpha \sim 0.01$ on the disk viscosity. A later study by \citet{King07} using AGN variability triggered by viscosity fluctuations suggest that $\alpha \sim 0.1$-$0.4$. \citet{Wood01} examined a variant of the \citet{Lyubarskii97} propagating fluctuations model in which episodic mass deposition at the outer edge of the accretion disk triggers viscosity fluctuations in the accretion disk. This model was revisited by \citet{Titarchuk07} who generalized the treatment of \citet{Wood01} to include time-dependent fluctuation sources located randomly on the disk face. \citet{Titarchuk07} follow the evolution of perturbations to the mass surface density and luminosity arising from viscosity that is a power-law function of disk radius. Subsequently, \citet{Kelly11} studied solutions to the surface mass density equation perturbed by stochastic fluctuations and showed that the resulting light curve can be expressed as a linear combination of multiple Ornstein-Uhlenbeck processes \citep{Gillespie96,Kelly09}.

Accretion disk variability may also be caused by the existence of `hot-spots' in the accretion disk (e.g. \citet{MaccaroneAccretion}). Hot spots are thought to arise when accretion streams impact the disk resulting in shock heating. Such hot spots have already been detected during periods of quiescence in BHXRBs \citep{Froning11,McClintock95}.

Thermal viscous instabilities or a variable mass transfer rate may be responsible for accretion disk variability \citep{Lasota01,Coriat12}, though the amplitude of the variability should be much smaller and the time-scale much longer for AGN than for BHXRBs \citep{Hameury09}.

Other sources of variability include magnetic flares in the corona of the accretion disk \citep{PoutanenFabian99} or from the hot accretion flow itself \citep{Veledina13}. Alternatively, dynamo processes due to the magnetic fields threading the disk may be responsible for producing variability \citep{LivioPringleKing03,King04,MayerPringle06}. \citet{JaniukCzerny07} have suggested that the variability may be produced in the hot X-ray corona surrounding the accretion disk from the evaporation of mass perturbations in the disk. \citet{Misra08} suggested that the X-ray variability of Cyg X-1 can be modeled as a damped harmonic oscillator perturbed by noise. Quasi-periodic signals in the light curve may be produced by weak shocks in the innermost region of the accretion caused by the presence of a tilt in the accretion disk of the black hole \citep{FragileBlaes07,FragileBlaes08}.

Quasi-periodic signals may be caused by global oscillation modes or parametric resonance instability \citep{ReynoldsMiller09a,ReynoldsMiller09b,Oneill11}. Variability may also be produced by standing shocks interacting with fluctuations in tilted accretion disks \citep{Henisey12}.

@@

Several mechanisms may be responsible for the variability observed in AGN light curves. Light curves in the optical are thought to be driven by X-ray variability on shorter time-scales but exhibit a local source of variability on longer time-scales \citep{UttleyAccretion}. This implies that it is unlikely that a single mechanism drives variability across the entire electromagnetic spectrum. Rather, the nature of the underlying mechanism responsible for variability may be different at different wavelengths.

@@

We propose a Green's function-based method for analysing observed light curves, particularly in the UV through IR where the fluctuations are smaller than they tend to be in the X-ray. Such a method facilitates comparison with analytic and simulation work on variability. Our method builds on the C-ARMA modelling technique of \citet{Kelly14} by proposing that the C-ARMA process is a direct result of the physics driving the variability.

It is likely that variability originates from multiple mechanisms simultaneously. These mechanisms are almost certainly non-linear \citep{UMV05}. We begin by linearising a general non-linear differential equation such as the surface mass density equation of~\eqref{eq:SurfaceMassDensity} that governs how perturbations evolve. We shall find that the linearisation of the non-linear equation results in a linear differential equation. This linearised equation has the form of the left hand side (LHS) of the C-ARMA process of equation~\eqref{eq:CARMAIntro}. Since the C-ARMA process is linear, we suggest that the Green's function of the LHS of equation~\eqref{eq:CARMAIntro} be used to probe dissipation physics in the accretion disk.

The right hand side (RHS) of the C-ARMA process of equation~\eqref{eq:CARMAIntro} can then be viewed as generating perturbations to the LHS, i.e. a C-ARMA process is a linear differential equation perturbed/driven by a random process. This driving process may be viewed as disturbances generated by MRI instabilities in the accretion disk. Section~\ref{sec:RHS} proposes a form for the disturbances driving the flux perturbations: coloured noise stochastic processes. A simple example of a C-ARMA process--the C-ARMA($2$,$1$) process--is presented in section~\ref{sec:CARMA21}.

@@

We assume that the variations in AGN luminosity may be ascribed to flux perturbations intrinsic to the accretion disk, corona and winds and that the total flux emitted by the AGN evolves via a (probably non-linear) differential equation. We shall examine how linear perturbations of the total flux evolve. Let the total flux emitted by an AGN be $F(t)$. Suppose $F$ obeys the $p^{\mathrm{th}}$-order non-linear differential equation 

@@

where $\mathrm{d}u$ is an increment of $u$. \textit{We see that the LHS of this equation is identical to the LHS of the C-ARMA process equation~\eqref{eq:CARMAIntro} suggesting that a C-ARMA process may be a good model for small flux variations.}

@@

We find that if the total flux emitted by the accretion disk obeys a non-linear differential equation, we can linearize that equation to determine the behaviour of small perturbations of the flux from the steady-state solution. The linear differential equation that governs how a small flux perturbation evolves is identical to the C-ARMA process equation suggesting that such processes may be good models for the stochastic variability seen in AGN accretion disks. In the next section, we present a convenient physical interpretation of equation~\eqref{eq:FinalSystemLHS} using the method of the Green's function.

@@

We compute the Green's function of equation~\eqref{eq:FinalSystemLHS} by finding a solution of the homogenized version of the equation. Consider the homogeneous differential equation corresponding to equation~\eqref{eq:FinalSystemLHS}
\begin{equation}\label{eq:HomoCARMA}
\frac{\mathrm{d}^{p}f}{\mathrm{d}t^{p}} + \alpha_{1} \frac{\mathrm{d}^{p-1}f}{\mathrm{d}t^{p-1}} + \ldots + \alpha_{p-1} \frac{\mathrm{d}f}{\mathrm{d}t} + \alpha_{p} f = 0.
\end{equation}
To obtain the Green's function solution of this equation, we shall drive this equation with a unit impulse, i.e. a Dirac delta function, located at $t = 0$
\begin{equation}\label{eq:GreenCARMA}
\frac{\mathrm{d}^{p}G}{\mathrm{d}t^{p}} + \alpha_{1} \frac{\mathrm{d}^{p-1}G}{\mathrm{d}t^{p-1}} + \ldots + \alpha_{p-1} \frac{\mathrm{d}G}{\mathrm{d}t} + \alpha_{p} G = \delta(t),
\end{equation}
where $G(t)$ is the desired Green's function solution. To find this solution, we shall first solve equation~\eqref{eq:HomoCARMA} by finding the roots of the characteristic polynomial of equation~\eqref{eq:ARCharPoly} and then apply boundary conditions based on the properties of $\delta(x)$. From the definition of the characteristic polynomial in equation~\eqref{eq:ARCharPoly}, the (distinct) roots are $\rho_{k}$ with $\operatorname{Re}(\rho_{k}) < 0$ for stability. Then the solution of the homogeneous equation~\eqref{eq:HomoCARMA} is 
\begin{equation}\label{eq:HomoSol}
f(t) = \sum_{k = 1}^{n} c_{k} \mathrm{e}^{\rho_{k}t},
\end{equation}
where the $c_{k}$ are constants. Hence the solution of the inhomogeneous equation~\eqref{eq:GreenCARMA} may be found by using the continuity properties of $G(t)$ and its derivatives. For $t < 0$, $\mathrm{d}^{p-k}G/\mathrm{d}t^{p-k} = 0$ for $1 \leq k < p$ while at $t = 0$, $\mathrm{d}^{p}G/\mathrm{d}t^{p}$ has the same type of discontinuity as $\delta(t)$. Therefore $\mathrm{d}^{p-1}G/\mathrm{d}t^{p-1} = 1$ when $t = 0$, i.e. it behaves like the step function. Next, we have that $\mathrm{d}^{p-k}G/\mathrm{d}t^{p-k} = 0$ at $t = 0$ for all $1 < k \leq p$ since each successive lower-order derivative is a $(k-1)^{\mathrm{th}}$-order polynomial located at the origin. Following this line of reasoning, we arrive at the following boundary conditions
\begin{equation}\label{eq:BC1a}
 \left.\frac{\mathrm{d}^{p-1}G}{\mathrm{d}t^{p-1}}\right|_{t = 0} = 1,\\
\end{equation}
and
\begin{equation}\label{eq:BC2a}
 \left.\frac{\mathrm{d}^{p-k}G}{\mathrm{d}t^{p-k}}\right|_{t = 0} = 0,
\end{equation}
for all $2 \leq k \leq p$. Computation of the derivatives in equations~\eqref{eq:BC1a} and \eqref{eq:BC2a} results in
\begin{equation}\label{eq:BC1b}
 \sum_{k = 1}^{p} c_{k} \rho_{k}^{p-1} = 1,
 \end{equation}
 and
\begin{equation}\label{eq:BC2b}
 \sum_{k = 1}^{p} c_{k} \rho_{k}^{k-1} = 0,
\end{equation}
for all $2 \leq k \leq p$. The $c_{k}$ may then be found by solving the matrix equation
\begin{equation}\label{eq:BCs}
\left( \begin{array}{ccccc} 
\rho_{1}^{p-1} & \rho_{2}^{p-1} & \hdots & \rho_{p-1}^{p-1} & \rho_{p}^{p-1} \\
\rho_{1}^{p-2} & \rho_{2}^{p-2} & \hdots & \rho_{p-1}^{p-2} & \rho_{p}^{p-2} \\
\vdots & \vdots & \ddots & \vdots & \vdots \\
\rho_{1} & \rho_{2} & \hdots & \rho_{p-1} & \rho_{p} \\
1 & 1 & \hdots & 1 & 1 \\
\end{array}\right) \left( \begin{array}{c} c_{1} \\ c_{2} \\ \vdots \\ c_{p-1} \\ c_{p} \end{array} \right) = \left( \begin{array}{c} 1 \\ 0 \\ \vdots \\ 0 \\ 0 \end{array} \right).
\end{equation}
In short, the Green's function for equation~\eqref{eq:FinalSystemLHS} is given by 
\begin{equation}\label{eq:GreenSol}
G(t) = \sum_{k = 1}^{p} c_{k} \mathrm{e}^{\rho_{k}t},
\end{equation}
where the $c_{k}$ satisfy equation~\eqref{eq:BCs}.

\section[Driving Perturbations]{How to Describe the Driver of the Flux Perturbations?}\label{sec:RHS}

@@

%\section[Fitting C-ARMA Processes]{How to Fit a C-ARMA Process to Observations?}\label{sec:Fitting}

%We wish to infer the values of the co-efficients in equation~\eqref{eq:CARMA} that defines C-ARMA  in order to compute the physically interesting Green's function and driving disturbance PSD for the observed light curve. While many techniques can be applied to perform this task, we use the Kalman filter \citep{Kalman60, Simon} since it requires trivial matrix inversions and determinant calculations ($1 \times 1$ matrices in the case of single-filter observations). An added advantage of the Kalman filter is that it is designed to allow us to update the likelihood calculation when new measurements become available. Finally, the framework in which the Kalman filter operates, the state-space framework, is also easy to extend to the case of multi-band, non-simultaneous observations of AGN as will become available with next generation surveys such as PanSTARRS and the LSST.

%The most straightforward method of fitting a C-ARMA process to a light curve is by direct estimation of the PSD of the observed light curve.
%Using periodogram estimators of the PSD, the observed PSD can be compared to the PSD of equation~\eqref{eq:CARMAPSD} which is of the form of a rational function of $\nu$. This rational function form for the C-ARMA PSD allows it great flexibility in matching any observed PSD. This is the method advocated in \citet*{Uttley02}.
